Informative Oceanographic Sampling using Excursion Probabilities for Multivariate Random Fields

### 2019-05-29 main_new.tex
updates

The excursions could also be defined as larger than threshold, or within a boundary of limits for the two variables. Calculations and interpretations will be similar in these different cases.

The ES is defined by random indicator variables, and the associated EPs become 
\begin{equation}\label{eq:prob}
 p(\bx) = P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b), \hspace{3mm} \bx \in \mathcal{M}.
\end{equation}
The actual calculation of the EP would require a model specification. As we demonstrate below, Gaussian modeling assumptions facilitate the computation. 
If the EPs in Eq. (\ref{eq:prob}) are close to $1$ or $0$, it is easy to classify water masses. It is more difficult if the probabilities are close to $0.5$. In this case it will likely be worthwhile gathering data to improve the classification skill.

There has lately been much statistical literature on ESs and EPs connected to spatial statistics \citep{french2013spatio,bolin2015excursion,french2016credible}. 
Most of the work done is based on GPs for the spatial phenomenon, as we use here.
Our focus in the current paper is on ES as defined in Eq. (\ref{ES}) and location-wise EPs and the uncertainty reduction that can be achieved by informative sampling. 
In this sense our work is similar to \cite{chevalier2014fast} and \cite{azzimonti2016quantifying} who described analytical results for uncertainty reduction in ESs for univariate processes.
One contribution of the current paper is to derive closed-form results for expectations of ESs in the situation with multivariate processes.

\section{Gaussian processes and excursion probabilities}\label{sec:GP_EP}

@@

We represent the bivariate random field of ocean temperature and salinity by a Gaussian model.
At location $\bx \in \mathcal{M}$, the bivariate distribution is 
\begin{equation}\label{gp_bar}
\begin{bmatrix}\xi_a(\bx) \\
\xi_b(\bx) 
\end{bmatrix}
 \sim N \left( 
\begin{bmatrix} \mu_{a}(\bx)\\
\mu_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
\sigma_{a}^2(\bx) & \sigma_{a}(\bx) \sigma_b(\bx) \gamma  \\
\sigma_{a}(\bx) \sigma_b(\bx) \gamma  & \sigma_{b}^2(\bx) 
\end{bmatrix}
\right).
\end{equation}
Here, the means and standard deviations are allowed to vary with location, which can be important for capturing the trends and variabilities near a river mouth. The correlation in salinity and temperature is assumed to be the same for all locations.

>>>>

 The excursions could also be defined as larger than a threshold, or
 within a boundary of limits for the two variables. Calculations and
 interpretations will be similar in these different cases.

The ES is defined by random indicator variables, and the associated EPs become 
\begin{equation}\label{eq:prob}
 p(\bx) = P(\xi_a(\bx) \leq T_a, \xi_b(\bx) \leq T_b), \hspace{3mm} \bx \in \mathcal{M}.
\end{equation}

The actual calculation of the EP would require a model
specification. As we demonstrate below, Gaussian modeling assumptions
facilitate the computation.  If the EPs in Eq. (\ref{eq:prob}) are
close to $1$ or $0$, it is easy to classify water masses; it is
challenging if the probabilities are close to $0.5$. In such a case it
would be worthwhile to gather data to improve classification skill.

Recent work in statistical literature on ESs and EPs connected to
spatial statistics includes
\citep{french2013spatio,bolin2015excursion,french2016credible}.  Most
are based on GPs for spatial phenomena, as in our work.  Our own focus
is on ES as defined in Eq. (\ref{ES}) and location-wise EPs and the
uncertainty reduction achieved by informative sampling.  In this sense
our work is similar to \cite{chevalier2014fast} and
\cite{azzimonti2016quantifying} who describe analytical results for
uncertainty reduction in ESs for univariate processes.  A major
contribution of our work is to derive closed-form results for
expectations of ESs in the situation with multivariate processes.

\section{Gaussian processes and excursion probabilities}
\label{sec:GP_EP}

@@

We represent the bivariate random field of ocean temperature and
salinity by a Gaussian model.  At location $\bx \in \mathcal{M}$, the
bivariate distribution is

\begin{equation}\label{gp_bar}
\begin{bmatrix}\xi_a(\bx) \\
\xi_b(\bx) 
\end{bmatrix}
 \sim N \left( 
\begin{bmatrix} \mu_{a}(\bx)\\
\mu_{b}(\bx)
\end{bmatrix},\begin{bmatrix}
\sigma_{a}^2(\bx) & \sigma_{a}(\bx) \sigma_b(\bx) \gamma  \\
\sigma_{a}(\bx) \sigma_b(\bx) \gamma  & \sigma_{b}^2(\bx) 
\end{bmatrix}
\right).
\end{equation}
Here, the means and standard deviations are allowed to vary with location, which can be important for capturing the trends and variabilities near a river mouth. The correlation in salinity and temperature is assumed to be the same for all locations.
